VFTS682: a confirmed dynamical ejection?

### 2018-05-09 VFTS682.tex
./gitignore

The ``R136 cluster'' is effectively defined by taking all the stars
within 25 arcseconds from R136a, one of the most massive members of
the cluster itself \citep[][]{crowther:10}, requiring the same
``quality'' criteria applied above. This selection yields 62 stars.
\todo{improve definition
  of stars from R136, describe accordingly -- check with Danny}

@@

Using the 437 stars selected as the ``surrounding population'' in
\Secref{sec:sample} and shown in blue in \Figref{fig:main}, we find averaged proper motion components of
$\langle\mu_\mathrm{RA}\rangle = 1.695\pm0.003\,\mathrm{mas\ yr^{-1}}$ and
$\langle\mu_\mathrm{DE}\rangle = 0.691\pm0.003\,\mathrm{mas\ yr^{-1}}$. Subtracting these values from the
proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\mu^\mathrm{sur}_\mathrm{RA} = 0.15\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
0.09\pm 0.08\,\mathrm{mas\ yr^{-1}}$. %% note that the error budget is
% dominated by VFTS682
These can be converted in the
the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=35\pm17\,\kms$,
$v^\mathrm{sur}_\mathrm{DE}=22\pm19\,\kms$, assuming a distance of
50\,kpc (we do not account for the uncertainty in the distance
estimate when propagating errors). The radial velocity from
\cite{bestenlehner:11} then gives the third component along
the line of sight, allowing us to calculate the speed of the star:

\begin{equation}
  \label{eq:speed}
  v = \sqrt{\left(v^\mathrm{sur}_\mathrm{RA}\right)^2
    +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 61 \pm 22
  \, \kms \ .
\end{equation}
This value for the three-dimensional speed of VFTS682 with respect the
surrounding stars make it a ``bona fide'' runaway star.

@@

The inset plot in \Figref{fig:main} shows a zoom in of the region
around R136 which includes VFTS682. To test the hypothesis that the
star was indeed ejected from this region, we check the
orientation of its proper motion relative to the group of stars
selected to represent the R136 cluster in \Secref{sec:sample}. We
further consider also the kinematic age of the star. The error cone, obtained by
propagating the errors on the proper motion components indicates that
R136 is indeed the most likely origin, as expected by \cite{fujii:11,
  banerjee:12}.

>>>>

The ``R136 cluster'' is effectively defined by taking all the stars
in a region of 50 arcseconds around R136a, one of the most massive members of
the cluster itself \citep[][]{crowther:10}, requiring the same
``quality'' criteria applied above. This selection yields 42 stars.
\todo{improve definition of stars from R136, describe accordingly -- check with Danny}

@@

Using the 437 stars selected as the ``surrounding population'' in
\Secref{sec:sample} and shown in blue in \Figref{fig:main}, we find averaged proper motion components of
$\langle\mu_\mathrm{RA}^\mathrm{sur}\rangle = 1.695\pm0.003\,\mathrm{mas\ yr^{-1}}$ and
$\langle\mu_\mathrm{DE}^\mathrm{sur}\rangle = 0.691\pm0.003\,\mathrm{mas\ yr^{-1}}$. Subtracting these values from the
proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\mu^\mathrm{sur}_\mathrm{RA} = 0.15\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
0.09\pm 0.08\,\mathrm{mas\ yr^{-1}}$. %% note that the error budget is
% dominated by VFTS682
These can be converted in the
the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=35\pm17\,\kms$,
$v^\mathrm{sur}_\mathrm{DE}=22\pm19\,\kms$, assuming a distance of
50\,kpc (we do not account for the uncertainty in the distance
estimate when propagating errors). The radial velocity from
\cite{bestenlehner:11} then gives the third component along
the line of sight, allowing us to calculate the speed of the star:

\begin{equation}
  \label{eq:speed_around}
  v^\mathrm{sur} = \sqrt{\left(v^\mathrm{sur}_\mathrm{RA}\right)^2
    +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 61 \pm 22
  \, \kms \ .
\end{equation}
This value for the three-dimensional speed of VFTS682 with respect the
surrounding stars make it a ``bona fide'' runaway star.

@@

The inset plot in \Figref{fig:main} shows a zoom in of the region
around R136 which includes VFTS682. Stars belonging to our ``R136
cluster'' group are shown in green (see \Secref{sec:sample}), together with their proper motion
as listed in the Gaia DR2 catalog.
To test the hypothesis that VFTS682 was indeed ejected from this region, we check the
orientation of its proper motion relative to these stars. We further
consider also the kinematic age of the star.

We begin by evaluating the average proper motion for the stars
representing the cluster. This is necessary to constrain the actual velocity
obtained by VFTS682 at its ejection. By using \Eqref{eq:mean} we find  
$\langle\mu_\mathrm{RA}^{R136}\rangle = 1.750\pm0.008\,\mathrm{mas\ yr^{-1}}$
and $\langle\mu_\mathrm{DE}^\mathrm{R136}\rangle =
0.742\pm0.009\,\mathrm{mas\ yr^{-1}}$. Therefore, using the values
from \Tabref{tab:vfts682}, we obtain the components of the proper
motion for VFTS682 relative to R136 $\mu^{R136}_\mathrm{RA} = 0.09\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^{R136}_\mathrm{DE} =
0.04\pm 0.08\,\mathrm{mas\ yr^{-1}}$.\todo{improve numbers by
  selecting better sample}.
\todo{discuss if direction is compatible
The error cone, obtained by
propagating the errors on the proper motion components indicates that
R136 is indeed the most likely origin, as expected by \cite{fujii:11,
  banerjee:12}.
}

Assuming again a distance of 50\,kpc and
not considering the error on the distance, we can convert these into
the components of the tangential velocity relative to R136,
$v^\mathrm{R136}_\mathrm{RA}=21 \pm 18\, \kms$ and $v^\mathrm{R136}_\mathrm{DE}=10
\pm 21\,\kms$. Finally, using the peculiar line of sight velocity
from \cite{bestenlehner:11}, we obtain the three-dimensional peculiar
velocity of VFTS682 with respect to R136

\begin{equation}
  \label{eq:speed_R136}
  v^\mathrm{sur} = \sqrt{\left(v^{R136}_\mathrm{RA}\right)^2
    +\left(v^{R136}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 51 \pm 23
  \, \kms \ .
\end{equation}
